Sperm should evolve to make female meiosis fair.

### 2014-07-26 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
small final changes to main text

\yb{Because female meiotic drive is likely common and consequential, 
	and because sperm have ample opportunity to influence female drive,  
	we develop population genetic models exploring the evolution of sperm influence on female meiotic drive -- 
	that is, what is the expected influence of sperm on female drive? } 
We first focus on models in which `self-promoting' alleles in
  sperm facilitate drive of like alleles during gametogenesis in heterozygous females. 
These models show that such sperm-acting alleles 
	have more difficulty invading a population than do traditional meiotic drivers, 
	 and under most circumstances, cannot be maintained as a balanced polymorphism.
Because self-promoting drivers are unlikely to create a sustained genomic conflict, 
	female meiosis will have little incentive to evolve resistance to them.
We then examine models in which a novel sperm-acting allele modifies the efficacy of a polymorphic meiotic driver. 
Such models universally reveal that sperm-acting drive modifiers are favored only if they suppress drive. 
These results highlight the fact that the interests of sperm and
maternal  genomes' are often aligned, 
as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
Thus, there is little selective benefit to females in preventing sperm to influence female meioses. 
\gc{Given the wide-spread requirement of fertilization for the completion
of  female meiosis it seems possible that various feature of the
interaction between sperm and egg have evolved as another mechanism to enforce an equitable transfer of genetic material.}
\gc{HOW ABOUT THE ABOVE.}

@@

\gc{Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
	for female meiotic drive and the requirement of fertilization for the completion of female meiosis. 
For example, we predict that a lower opportunity for female meiotic drive,
        e.g. an animal lineage with a history of high inbreeding
        or selfing, 
        may be accompanied by a relaxation of the requirement of fertilization for the
        completion of female meiosis (although 
   \yb{     opportunities to test this hypothesis may be limited because }
        %this effect may be weaks as such
        lineages may only persist for a short time). 
Our results also suggests that 
%Another   hypothesis that our work suggests is we might expect 
        that phylogenetic variation in the stage of female meiosis when
        fertilization occurs may influence the prevalence of
        female meiotic drive.
For example, in taxa where meiosis I is completed
by the egg pre-fertilization, including humans, centromere drive may
be more common as sperm based modifiers can not intercede than in
species where fertilization occurs during MI. 
Such hypotheses could be tested if the speed of centromere
turnover serves as a proxy for the efficacy of MI drive. }
\newline

@@

Our results highlight potentially counterintuitive results of complex genetic conflicts. 
Despite much opportunity for conflict between sperm and females over fertilization \citep{Partridge1998},
	the interests of fertilizing sperm and female are quite well aligned during syngamy.
While conflict between mother and her alternative chromosomes ensues, 
	fertilizing sperm decidedly side with mom, as both have a shared interest in producing a viable and potentially fit offspring. 
\yb{Our model does not directly speak to the evolutionary origin of
  female meiotic arrest \citep[for a review and evaluation of such
  hypotheses see][]{Mira1998}.
However, given the existence of female meiotic arrest, 
	and that its timing relative to
	fertilization as well as the interactions between sperm and egg are variable across species  \citep{Karrbook:09} 
	the nature of the meiotic arrest and interactions between sperm and egg may be molded by selection to reduce the opportunity for female meiotic drive, 
	while selfish drivers may evolve in concert, to overcome these adaptations. 
%leave female meiosis open to sabotage by sperm collaborators, but
%rather that  an 
%evolutionary loss of female meiotic arrest will result in the loss of sperm defense against female meiotic drivers. } 
  }
%However, given that arrest does occur its timing relative to
%fertilization, various features of the interaction between sperm and
%egg and divergence of these features across species  \citep{Karrbook:09} could be subject
%to selection to conflict between female meiotic drivers and alleles
%elsewhere in the genomes of sperm and eggs.
%the mechanism by which it occur
%mechanistic features, and evolutionary changes among taxa
%various features of the
%mechanism of fertilization 
%however, our work does imply that the fertilization requirement does
%not 
%leave female meiosis open to sabotage by sperm collaborators, but
%rather that  an 
%evolutionary loss of female meiotic arrest will result in the loss of sperm defense against female meiotic drivers. }

>>>>

\yb{Because female meiotic drive is likely common and consequential, 
	and because sperm have ample opportunity to influence female drive,  
	we develop population genetic models exploring the evolution of sperm influence on female meiotic drive -- 
	that is, what is the expected influence of sperm on female drive? } 
We first focus on models in which `self-promoting' alleles in
  sperm facilitate drive of like alleles during gametogenesis in heterozygous females. 
These models show that such sperm-acting alleles 
	have more difficulty invading a population than do traditional meiotic drivers, 
	 and under most circumstances, cannot be maintained as a balanced polymorphism.
Because self-promoting drivers are unlikely to create a sustained genomic conflict, 
	female meiosis will have little incentive to evolve resistance to them.
We then examine models in which a novel sperm-acting allele modifies the efficacy of a polymorphic meiotic driver. 
Such models universally reveal that sperm-acting drive modifiers are favored only if they suppress drive. 
These results highlight the fact that the interests of sperm and
maternal  genomes' are often aligned, 
as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
Thus, there is little selective benefit to females in preventing sperm to influence female meioses. 
\gc{Given the wide-spread requirement of fertilization for the completion
of  female meiosis it seems possible that various features of the
interaction between sperm and egg have evolved as a mechanism to enforce an equitable transfer of genetic material.}

@@

\gc{Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
	for female meiotic drive and the requirement of fertilization for the completion of female meiosis. 
For example, we predict that a lower opportunity for female meiotic drive,
        e.g. an animal lineage with a history of high inbreeding
        or selfing, 
        may be accompanied by a relaxation of the requirement of fertilization for the
        completion of female meiosis (although 
        opportunities to test this hypothesis may be limited because 
        %this effect may be weaks as such
        lineages may only persist for a short time). 
Our results also suggests that 
%Another   hypothesis that our work suggests is we might expect 
        that phylogenetic variation in the stage of female meiosis when
        fertilization occurs may influence the prevalence of
        female meiotic drive.
\gc{For example, in taxa where meiosis I is completed
by the egg pre-fertilization centromere drive may
be more common as sperm-based modifiers can not intercede, as compared to
species where fertilization occurs during MI. }
Such hypotheses could be tested if the speed of centromere
turnover serves as a proxy for the efficacy of MI drive. 
\newline

@@

Our results highlight potentially counterintuitive results of complex genetic conflicts. 
Despite much opportunity for conflict between sperm and females over fertilization \citep{Partridge1998},
	the interests of fertilizing sperm and female are quite well aligned during syngamy.
While conflict between mother and her alternative chromosomes ensues, 
	fertilizing sperm decidedly side with mom, as both have a shared interest in producing a viable and potentially fit offspring. 
Our model does not directly speak to the evolutionary origin of
  female meiotic arrest \citep[for a review and evaluation of such
  hypotheses see][]{Mira1998}.
However, given the existence of female meiotic arrest, 
	\gc{and that its timing and mechanistic details are variable across species \citep{Masui_book,Karrbook:09} 
	the nature of the meiotic arrest and interactions between
        sperm and egg may  be molded by selection to reduce the opportunity for female meiotic drive, 
	and selfish drivers evolving in concert to overcome these adaptations. }
%leave female meiosis open to sabotage by sperm collaborators, but
%rather that  an 
%evolutionary loss of female meiotic arrest will result in the loss of sperm defense against female meiotic drivers. }

%However, given that arrest does occur its timing relative to
%fertilization, various features of the interaction between sperm and
%egg and divergence of these features across species  \citep{Karrbook:09} could be subject
%to selection to conflict between female meiotic drivers and alleles
%elsewhere in the genomes of sperm and eggs.
%the mechanism by which it occur
%mechanistic features, and evolutionary changes among taxa
%various features of the
%mechanism of fertilization 
%however, our work does imply that the fertilization requirement does
%not 
%leave female meiosis open to sabotage by sperm collaborators, but
%rather that  an 
%evolutionary loss of female meiotic arrest will result in the loss of sperm defense against female meiotic drivers. }
